### 2016-02-20 ch03-lrm/LPMSmooth_ML_EG_JS_JL.tex
Clean up chapter 3

% \title{
% Smoothing the LPM-Estimate of the Frequency Response Function Via an Impulse Response Truncation Technique}

% \author{$^1$M.L.D. Lumori, \emph{Senior Member, IEEE}, $^2$E. Geerardyn, \emph{Student Member, IEEE}, \\ $^2$J. Schoukens, \emph{Fellow, IEEE}, and
% $^2$J. Lataire, \emph{Member, IEEE}\\

% \thanks{$^1$University of San Diego, Department of Electrical Engineering, 5998 Alcal\'a Park, San Diego, CA 92110, USA {\tt\small mlumori@sandiego.edu}}
% \thanks{$^2$Vrije Universiteit Brussel, Dept. ELEC, Fundamental Electricity and Instrumentation, Pleinlaan 2, 1050 Brussels, Belgium}
% \thanks{This work was supported in part by the Fund for Scientific Research (FWO-Vlaanderen), by the Flemish Government (Methusalem Fund, METH1), and by the Belgian Federal Government (IAP VII DYSCO).}
% }

% \begin{document}
% \maketitle

% \begin{abstract}
% A statistical impulse response truncation technique is applied to the local-polynomial-method-estimate of the frequency response function, resulting in an improved, smooth frequency response function. Formulated as a nonparametric linear-least-squares-estimate, the local polynomial method is first applied to estimate the frequency response function from a full data record of a single-input-single-output system, systematically expressed in an output-error framework. The smooth characteristics of both the exact frequency response function and the leakage from transients allow for an optimal application of the local polynomial method, leading to the elimination of both the leakage and interpolation errors. The truncation method introduced in this paper makes it possible for the user to fine-tune the trade-off between the uncertainty (variance) and the bias on the estimated instantaneous frequency response function.
% \end{abstract}

% \begin{IEEEkeywords}
% Smoothing methods, Estimation, Least squares methods, Frequency response, Frequency domain analysis
% \end{IEEEkeywords}
\researchBasedOn[This section]{Lumori2014TIM}

@@

\TODO{remove duplicate labels (global)}

@@

This paper uses nonparametric models in conjunction with the \gls{LPM}~\citep{Pintelon2010LPM1} to estimate the complex-valued \gls{FRF} in a linear \gls{LS} fashion.
The advantage of using the nonparametric models is basically the non-involvement of the user in choosing a model order, i.e. no user interaction. Among the current nonparametric methods, the LPM method is chosen due to its superiority over the classical nonparametric methods (windowing) in suppressing leakage errors \citep{Bendat1993,Oppenheim1983}. Consequently, the LPM estimate of the FRF is not hampered by leakage errors \citep{Pintelon2010LPM1,Schoukens2009LPM}. Note that a more general discussion on the LPM, applied to noisy, real-valued data is found in \citep{Fan1996} with a specific focus on parametric tuning.

The main contribution of the paper is to come up with an improved FRF by developing a method for smoothing the FRF estimated via the LPM. This is done by truncating the associated impulse response, as in~\citep{Schoukens1998}, but with the following additional extension: the determination of the optimal truncation time, without any user interaction, in conjunction with the use of the LPM for leakage reduction. The cutoff index (truncation point) is determined by a statistically sound method. It is then shown that a smooth LPM estimate of the FRF lowers the variance, thus improving the assessment level of the dynamic properties of a given LTI system.

Based on \citep{Schoukens2009LPM}, the nonparametric method developed in this work is formulated in an output-error (OE) framework. A single-input-single-output (SISO) LTI system is considered. %, based on  \citep{Schoukens2009LPM}.
The input signal is a known random noise signal and the output signal is corrupted by measurement noise. This is depicted in \figref{lpmtdrep} as a linear dynamic discrete-time SISO system, whose full mathematical model is of the form:
\begin{equation}\label{eq:lpm:TD}
y(t)=G_0(q)u_0(t)+H_0(q)e(t)=G_0(q)u_0(t)+v(t)
\end{equation}
where $G_0(q)$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= H_0(q)e(t)$ is the noise source at the output, $H_0(q)$ is the noise dynamics, $e(t)$ is white noise, and $q^{-m}$ is the backwards shift operator ($q^{-m}x(t)$ = $x(t-m)$  with $m$ a positive integer). %\JL{$G_0(q)$ represents the dynamics of the system to be estimated.}

@@

In this paper, however, a nonparametric technique is developed, %the identification problem is
formulated in the frequency domain, consistent with \citep{Pintelon2012,Mahata2006}. The following  choices are made:

@@

The work in this paper is for discrete-time systems. Denote the $k^{th}$ discretized frequency as $\Omega_k$ = $e^{-j2{\pi}kf_s/N}$, with $f_s$ the sample frequency and $N$ the total number of sample points.
\item

Describe the parametric plant model  $G(q,\theta)$ by the nonparametric FRF counterpart  $G(\Omega_k)$  at the $k^{th}$ frequency.
\item

@@

The rest of the paper is structured as follows. \secref{se:theoryLPMandWindowing} covers the theory on the LPM and windowing for an OE setup.
\secref{se:smoothingFRFestimate} discusses the novel smoothing method.

Ultimately, in \secref{se:simResults}, simulation results of the following FRF estimates and their corresponding variances are compared: (i) the smooth LPM estimate $\hat{G}_\text{sm-poly}(\Omega_k)$, (ii) the LPM estimate $\hat{G}_\text{poly}(\Omega_k)$.

@@

Using convolution, the noiseless output signal in \figref{lpmtdrep} is $y_0(t) = g_0(t)*u_0(t)$, where the signals are sampled at $t = 0, 1, 2,...,N-1$ and the excitation signal $u_0(t)$ is random noise. As shown in \citep{Pintelon2012} and \citep{Pintelon1997} the \gls{DFT} spectra of the sampled signals at the frequency $\Omega_k$ are related as follows:
\begin{equation}\label{lpmleak}
Y_0(k)=G_0(\Omega_k)U_0(k)+T_G(\Omega_k)
\end{equation}
where $U_0(k)$ is the \gls{DFT} spectrum of the random noise excitation,  $G_0(\Omega_k)$ is the transfer function and $T_G(\Omega_k)$ is the transients leakage term.
In this paper, the \gls{DFT} of a signal $x(t)$ is defined explicitly as  \citep{Oppenheim1983}

@@

The following assumptions are consistent with \citep{Schoukens2009LPM} for an OE setup (\figref{lpmtdrep}):
\begin{assumption}
The SISO system $G_0$ and  the transients leakage spectrum $T_G$ are both smooth functions of the frequency.
\end{assumption}

@@

The spectrum of the  input signal $U_0$ is not a smooth function of the frequency, but a rough function, for example $U_0(k+1) - U_0(k)$ should not vanish to zero. %i.e. $U_0(k+1) - U_0(k) \neq 0$.
\end{assumption}

@@

The anomalies due to interpolation errors, associated with windowing, can easily be mitigated or circumvented by use of the LPM. As shown in \citep{Pintelon2012}, the leakage errors are reduced from  $\mathcal{O}({N}^{-1})$, for Hanning windowing, to $\mathcal{O}({N}^{-3})$ or better, for the LPM.

\paragraph{Formulation of the Linear LS LPM}
The LPM is formulated as a nonparametric local-linear-least-squares (LS) estimate using the full data record of length $N$ as outlined below \citep{Schoukens2009LPM}.
\citet[Section 7.2.2]{Pintelon2012} gives the general formulation of the \gls{LPM} for \gls{MIMO} systems.
The formulation in this paper is a summary for a \gls{SISO} system.

@@

The formulation of equation \eqref{eq:LSsolutionLPM} must take into account the fact that  a full rank of $K_n$ requires $n \geq R + 1$. The choice $n = R + 1$ results in the highest variance on the estimate, but with the smallest interpolation error \citep{Schoukens2009LPM}. %(per JL)!

@@

When substituted into equations \eqref{lpmGTaylorS} and \eqref{lpmGTMatrix},  these yield the following quadratic local polynomials of $G_0$ and $T$ around a central frequency $\Omega_{k}$:
\begin{align}
G_0(\Omega_{k+r})
  &\approx 
  G_0(\Omega_k)+g_1r+g_2r^2
  \label{eq:lpm:expansionG:quadratic}
\\
  T(\Omega_{k+r})
    &\approx 
    T(\Omega_k)+t_1r+t_2r^2
    \label{eq:lpm:expansionT:quadratic}
\end{align}
with ($r\in\mathbb{W}_3$) and the vector of parameters:
\begin{equation}\label{lpmThetaEst}
\theta^\top(k)=\left[G_0(\Omega_k) \ g_1 \ g_2\  T_G(k)\  t_1 \ t_2\right]
\end{equation}
The \gls{LPM} estimate of the FRF is the first element of the estimate $\hat\theta(k)$, obtained from \eqref{eq:LSsolutionLPM}, \emph{viz}:
\begin{equation}
\hat{G}_\text{poly}(\Omega_k) = \hat\theta_1(k)
\end{equation}

The estimation of $\hat{G}_\text{poly}(\Omega_k)$ at a single frequency $\Omega_k = 0.586$~Hz is illustrated in \figref{LPM_Schematic_EG}.% centered at $k$, where $r = 0$.
\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[scale=0.92]{\thisDir/figs/spectUOmegak}
   \includegraphics[scale=0.92]{\thisDir/figs/spectYOmegak}
   \includegraphics{\thisDir/figs/LPMfigOmegak}
   \caption{Illustration of the \gls{LPM}. Top figures: input spectrum (left) and output spectrum (right). A small band is selected (gray area, top figure), in which $\hat G_\mathrm{poly}(\Omega_{k+r})$ is estimated (black thick line, bottom figure) as a local polynomial. Only its central value $\hat G_\mathrm{poly}(\Omega_k)$ (white circle) is retained. This procedure is repeated for all $\Omega_k$.}
   \label{LPM_Schematic_EG}
\end{figure}
This procedure is repeated at each central frequency $\Omega_k$ for  $k\in \{1,\dots,N/2\}$ (i.e. the entire frequency band, including Nyquist, but excluding DC, and regularizing at the boundaries). This completes the \gls{LPM}, yielding an intrinsic averaging of the spectra of the estimated \gls{FRF}, $\hat{G}_\text{poly}(\Omega_k)$.

@@

\begin{enumerate}
\item Obtain the estimate of the impulse response $\hat g_\mathrm{poly}(t)$ from the LPM, corrected for its DC value as discussed in \secref{se:DCvalueEst}. Henceforth, $\hat g_\mathrm{poly}(t)$ will simply be denoted as $\hat g(t)$. %In the remainder of this procedure description $\hat g_\mathrm{poly}(t)$ will be simply denoted $\hat g(t)$.

@@

\item Obtain $\mathbb{T}_\mathrm{HSNR}$ as the set of time instants where $\hat g(t)$ is significantly above the standard deviation of the noise. Only samples with absolute values of at least $5\hat\sigma_\mathrm{n}$ are retained: %Only samples with an absolute value of at least $5\hat\sigma_\mathrm{n}$ have been retained:
\begin{align}
\mathbb{T}_\mathrm{HSNR} = \left\{
t:|\hat g(t)|\geqslant 5\hat\sigma_\mathrm{n}
\right\}.
\end{align}

\noindent
(Subscript HSNR stands for High \gls{SNR}.)

@@

\item
Determine the first time-instant $t_\mathrm{trunc}$ at which the estimated exponential gets significantly below the noise floor, \emph{viz}:

\begin{align}\label{eq:truncTimeExpFit}
t_\mathrm{trunc} = \min \left\{t:Ae^{at} < \gamma\hat\sigma_\mathrm{n}\right\}
\end{align}

\noindent
where the parameter $\gamma$ can be tuned such that an appropriate trade-off between the decreased variance and the increased bias on the estimated smoothed FRF is found. This is discussed below. %(see the discussion below).
% The value of $\gamma$ to be used depends on the system. %$\gamma = 0.1$ was found to be a good rule of thumb.

@@

The paper introduced a novel time domain method to smooth the \gls{LPM}-estimate of an \gls{FRF}. It consisted of, after obtaining the \gls{FRF} from the \gls{LPM}, computing the associated estimated impulse response via the \gls{IDFT}.
Then, it was determined statistically at which time index the impulse response had decayed below the noise floor, yielding a point beyond which the response may be set to zero.

>>>>

% \begin{abstract}
% A statistical impulse response truncation technique is applied to the local-polynomial-method-estimate of the frequency response function, resulting in an improved, smooth frequency response function. Formulated as a nonparametric linear-least-squares-estimate, the local polynomial method is first applied to estimate the frequency response function from a full data record of a single-input-single-output system, systematically expressed in an output-error framework. The smooth characteristics of both the exact frequency response function and the leakage from transients allow for an optimal application of the local polynomial method, leading to the elimination of both the leakage and interpolation errors. The truncation method introduced in this section makes it possible for the user to fine-tune the trade-off between the uncertainty (variance) and the bias on the estimated instantaneous frequency response function.
% \end{abstract}

\researchBasedOn[This section]{Lumori2014TIM}

@@

@@

This section uses nonparametric models in conjunction with the \gls{LPM}~\citep{Pintelon2010LPM1} to estimate the complex-valued \gls{FRF} in a linear \gls{LS} fashion.
The advantage of using the nonparametric models is basically the non-involvement of the user in choosing a model order, i.e. no user interaction. Among the current nonparametric methods, the \gls{LPM} method is chosen due to its superiority over the classical nonparametric methods (windowing) in suppressing leakage errors \citep{Bendat1993,Oppenheim1983}. 
Consequently, the \gls{LPM} estimate of the \gls{FRF} is not hampered by leakage errors \citep{Pintelon2010LPM1,Schoukens2009LPM}. Note that a more general discussion on the \gls{LPM}, applied to noisy, real-valued data is found in \citep{Fan1996} with a specific focus on parametric tuning.

The main contribution of this section is to come up with an improved \gls{FRF} by developing a method for smoothing the \gls{FRF} estimated via the LPM. This is done by truncating the associated impulse response, as in~\citep{Schoukens1998}, but with the following additional extension: the determination of the optimal truncation time, without any user interaction, in conjunction with the use of the \gls{LPM} for leakage reduction. The cutoff index (truncation point) is determined by a statistically sound method. It is then shown that a smooth \gls{LPM} estimate of the \gls{FRF} lowers the variance, thus improving the assessment level of the dynamic properties of a given LTI system.

Based on \citep{Schoukens2009LPM}, the nonparametric method developed in this work is formulated in an output-error (OE) framework. A \gls{SISO} \gls{LTI} system is considered. %, based on  \citep{Schoukens2009LPM}.
The input signal is a known random noise signal and the output signal is corrupted by measurement noise. This is depicted in \figref{lpmtdrep} as a linear dynamic discrete-time \gls{SISO} system, whose full mathematical model is of the form:
\begin{equation}\label{eq:lpm:TD}
y(t)=G_0(q)u_0(t)+H_0(q)e(t)=G_0(q)u_0(t)+v(t)
\end{equation}
where $G_0(q)$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= H_0(q)e(t)$ is the noise source at the output, $H_0(q)$ is the noise dynamics, $e(t)$ is white noise, and $q^{-m}$ is the backwards shift operator ($q^{-m}x(t)$ = $x(t-m)$  with $m$ a positive integer). %\JL{$G_0(q)$ represents the dynamics of the system to be estimated.}

@@

In this section, however, a nonparametric technique is developed, %the identification problem is
formulated in the frequency domain, consistent with \citep{Pintelon2012,Mahata2006}. The following  choices are made:

@@

The work in this section is for discrete-time systems. Denote the $k^{th}$ discretized frequency as $\Omega_k$ = $e^{-j2{\pi}kf_s/N}$, with $f_s$ the sample frequency and $N$ the total number of sample points.
\item

Describe the parametric plant model  $G(q,\theta)$ by the nonparametric \gls{FRF} counterpart  $G(\Omega_k)$  at the $k^{th}$ frequency.
\item

@@

The rest of the section is structured as follows. \secref{se:theoryLPMandWindowing} covers the theory on the \gls{LPM} and windowing for an \gls{OE} setup.
\secref{se:smoothingFRFestimate} discusses the novel smoothing method.

Ultimately, in \secref{se:simResults}, simulation results of the following FRF estimates and their corresponding variances are compared: (i) the smooth \gls{LPM} estimate $\hat{G}_\text{sm-poly}(\Omega_k)$, (ii) the \gls{LPM} estimate $\hat{G}_\text{poly}(\Omega_k)$.

@@

Using convolution, the noiseless output signal in \figref{lpmtdrep} is $y_0(t) = g_0(t)*u_0(t)$, where the signals are sampled at $t = 0, 1, 2,...,N-1$ and the excitation signal $u_0(t)$ is random noise. As shown in \citep{Pintelon2012} and \citep{Pintelon1997} the \gls{DFT} spectra of the sampled signals at the frequency $\Omega_k$ are related as follows:
\begin{equation}\label{lpmleak}
Y_0(k)=G_0(\Omega_k)U_0(k)+T_G(\Omega_k)
\end{equation}
where $U_0(k)$ is the \gls{DFT} spectrum of the random noise excitation,  $G_0(\Omega_k)$ is the transfer function and $T_G(\Omega_k)$ is the transients leakage term.
In this section, the \gls{DFT} of a signal $x(t)$ is defined explicitly as  \citep{Oppenheim1983}

@@

The following assumptions are consistent with \citep{Schoukens2009LPM} for an OE setup (\figref{lpmtdrep}):
\begin{assumption}
The \gls{SISO} system $G_0$ and  the transients leakage spectrum $T_G$ are both smooth functions of the frequency.
\end{assumption}

@@

The spectrum of the  input signal $U_0$ is not a smooth function of the frequency, but a rough function, for example $U_0(k+1) - U_0(k)$ should not vanish to zero. %i.e. $U_0(k+1) - U_0(k) \neq 0$.
\TODO{assumption is not complete! some expectation is missing!}
\end{assumption}

@@

The anomalies due to interpolation errors, associated with windowing, can easily be mitigated or circumvented by use of the \gls{LPM}. As shown in \citep{Pintelon2012}, the leakage errors are reduced from  $\mathcal{O}({N}^{-1})$, for Hanning windowing, to $\mathcal{O}({N}^{-3})$ or better, for the \gls{LPM}.

\paragraph{Formulation of the Linear LS LPM}
The \gls{LPM} is formulated as a nonparametric local-linear-least-squares (LS) estimate using the full data record of length $N$ as outlined below \citep{Schoukens2009LPM}.
\citet[Section 7.2.2]{Pintelon2012} gives the general formulation of the \gls{LPM} for \gls{MIMO} systems.
The formulation in this section is a summary for a \gls{SISO} system.

@@

The formulation of equation \eqref{eq:LSsolutionLPM} must take into account the fact that  a full rank of $K_n$ requires $n \geq R + 1$. The choice $n = R + 1$ results in the highest variance on the estimate, but with the smallest interpolation error \citep{Schoukens2009LPM}.

@@

When substituted into equations \eqref{lpmGTaylorS} and \eqref{lpmGTMatrix},  these yield the following quadratic local polynomials of $G_0$ and $T$ around a central frequency $\Omega_{k}$:
\begin{align}
G_0(\Omega_{k+r})
  &\approx 
  G_0(\Omega_k)+g_1r+g_2r^2
  \label{eq:lpm:expansionG:quadratic}
\\
  T(\Omega_{k+r})
    &\approx 
    T(\Omega_k)+t_1r+t_2r^2
    \label{eq:lpm:expansionT:quadratic}
\end{align}
with ($r\in\mathbb{W}_3$) and the vector of parameters:
\begin{equation}\label{lpmThetaEst}
\theta^\top(k)=\left[G_0(\Omega_k) \ g_1 \ g_2\  T_G(k)\  t_1 \ t_2\right]
\end{equation}
The \gls{LPM} estimate of the \gls{FRF} is the first element of the estimate $\hat\theta(k)$, obtained from \eqref{eq:LSsolutionLPM}, \emph{viz}:
\begin{equation}
\hat{G}_\text{poly}(\Omega_k) = \hat\theta_1(k)
\end{equation}

The estimation of $\hat{G}_\text{poly}(\Omega_k)$ at a single frequency $\Omega_k = 0.586\unit{Hz}$ is illustrated in \figref{LPM_Schematic_EG}.% centered at $k$, where $r = 0$.
\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[scale=0.92]{\thisDir/figs/spectUOmegak}
   \includegraphics[scale=0.92]{\thisDir/figs/spectYOmegak}
   \includegraphics{\thisDir/figs/LPMfigOmegak}
   \caption{Illustration of the \gls{LPM}. Top figures: input spectrum (left) and output spectrum (right). A small band is selected (gray area, top figure), in which $\hat G_\mathrm{poly}(\Omega_{k+r})$ is estimated (black thick line, bottom figure) as a local polynomial. Only its central value $\hat G_\mathrm{poly}(\Omega_k)$ (white circle) is retained. This procedure is repeated for all $\Omega_k$.}
   \label{LPM_Schematic_EG}
\end{figure}
This procedure is repeated at each central frequency $\Omega_k$ for  $k\in \Set{1,\dots,N/2}$ (i.e. the entire frequency band, including Nyquist, but excluding \gls{DC}, and regularizing at the boundaries). This completes the \gls{LPM}, yielding an intrinsic averaging of the spectra of the estimated \gls{FRF}, $\hat{G}_\text{poly}(\Omega_k)$.

@@

\begin{enumerate}
\item Obtain the estimate of the impulse response $\hat g_\mathrm{poly}(t)$ from the \gls{LPM}, corrected for its \gls{DC} value as discussed in \secref{se:DCvalueEst}. Henceforth, $\hat g_\mathrm{poly}(t)$ will simply be denoted as $\hat g(t)$. %In the remainder of this procedure description $\hat g_\mathrm{poly}(t)$ will be simply denoted $\hat g(t)$.

@@

\item Obtain $\mathbb{T}_\mathrm{HSNR}$ as the set of time instants where $\hat g(t)$ is significantly above the standard deviation of the noise. Only samples with absolute values of at least $5\hat\sigma_\mathrm{n}$ are retained: %Only samples with an absolute value of at least $5\hat\sigma_\mathrm{n}$ have been retained:
\begin{align}
\mathbb{T}_\mathrm{HSNR} = \left\{
t:|\hat g(t)|\geqslant 5\hat\sigma_\mathrm{n}
\right\}.
\end{align}
(Subscript HSNR stands for High \gls{SNR}.)

@@

\item
Determine the first time-instant $t_\mathrm{trunc}$ at which the estimated exponential gets significantly below the noise floor, \emph{viz}:
\begin{align}\label{eq:truncTimeExpFit}
t_\mathrm{trunc} = \min \left\{t:Ae^{at} < \gamma\hat\sigma_\mathrm{n}\right\}
\end{align}
where the parameter $\gamma$ can be tuned such that an appropriate trade-off between the decreased variance and the increased bias on the estimated smoothed FRF is found. This is discussed below. %(see the discussion below).
% The value of $\gamma$ to be used depends on the system. %$\gamma = 0.1$ was found to be a good rule of thumb.

@@

The section introduced a novel time domain method to smooth the \gls{LPM}-estimate of an \gls{FRF}. It consisted of, after obtaining the \gls{FRF} from the \gls{LPM}, computing the associated estimated impulse response via the \gls{IDFT}.
Then, it was determined statistically at which time index the impulse response had decayed below the noise floor, yielding a point beyond which the response may be set to zero.
